---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rsimsum

[![Travis-CI Build Status](https://travis-ci.org/ellessenne/rsimsum.svg?branch=master)](https://travis-ci.org/ellessenne/rsimsum)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ellessenne/rsimsum?branch=master&svg=true)](https://ci.appveyor.com/project/ellessenne/rsimsum)
[![Coverage Status](https://img.shields.io/codecov/c/github/ellessenne/rsimsum/master.svg)](https://codecov.io/github/ellessenne/rsimsum?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rsimsum)](https://cran.r-project.org/package=rsimsum)
[![CRAN_Logs_Badge](http://cranlogs.r-pkg.org/badges/rsimsum)](https://cran.r-project.org/package=rsimsum)

`rsimsum` is a porting of the user-written [simsum](http://www.stata-journal.com/article.html?article=st0200) command in Stata.

## Installation

You can install rsimsum from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ellessenne/rsimsum")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```

# References

* White, I.R. 2010. _simsum: Analyses of simulation studies including Monte Carlo error_. The Stata Journal 10(3): 369-385, [](http://www.stata-journal.com/article.html?article=st0200)
* Morris, T.P, White, I.R. and Crowther, M.J. 2017. _Using simulation studies to evaluate statistical methods_. <[arXiv:1712.03198](https://arxiv.org/abs/1712.03198)>
