---
title: "Nested loop plots"
author: "Alessandro Gasparini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nested loop plots}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
options(width = 150)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center", fig.width = 7,
  out.width = "80%"
)
```

As of version `0.6.0`, `rsimsum` supports the fully automated creation of nested loop plots (Rücker and Schwarzer, 2014).

```{r package}
library(rsimsum)
```

A dataset that can be purposefully used to illustrate nested loop plots is bundled and shipped with `rsimsum`:

```{r data}
data("nlp", package = "rsimsum")
```

This data set contains the results of a simulation study on survival modelling with 150 distinct data-generating mechanisms:

```{r display-data}
head(nlp)
```

Further information on the data could be find in the help file (`?nlp`).

We can analyse this simulation study using `rsimsum` as usual:

```{r analyse}
s <- rsimsum::simsum(
  data = nlp, estvarname = "b", true = 0, se = "se",
  methodvar = "model", by = c("baseline", "ss", "esigma")
)
s
```

Finally, a nested loop plot can be automatically produced via the `autoplot` method, e.g. for bias:

```{r nlp, fig.asp = 0.75}
library(ggplot2)
autoplot(s, type = "nlp", stats = "bias")
```

# References

* Rücker, G. and Schwarzer, G. 2014 _Presenting simulation results in a nested loop plot_. BMC Medical Research Methodology 14(1) <[doi:10.1186/1471-2288-14-129](https://doi.org/10.1186/1471-2288-14-129)>
