---
title: "Custom input values for confidence intervals and true values"
author: "Alessandro Gasparini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom input values for confidence intervals and true values}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
options(width = 150)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center", fig.height = 6, fig.width = 6,
  out.width = "75%"
)
```

`rsimsum` supports custom input values for the true value of the estimand and for confidence intervals limits (used to calculate coverage probability).

To illustrate this feature, we can use the `tt` dataset (bundled with `rsimsum`):

```{r packages}
library(rsimsum)
data("tt", package = "rsimsum")
head(tt)
```

This includes the results of a simulation study assessing robustness of the t-test when estimating the difference between means.
The t-test assumes a t distribution, hence confidence intervals for the estimated mean are generally based on the t distribution.
See for instance the example from the t-test documentation (`?t.test`):

```{r}
t.test(extra ~ group, data = sleep)
```

We can incorporate custom confidence intervals by passing the name of two columns in `data` as the `ci.limits` argument:

```{r}
s1 <- simsum(data = tt, estvarname = "diff", true = -1, se = "se", ci.limits = c("lower", "upper"), methodvar = "method", by = "dgm")
summary(s1, stats = "cover")
```

By doing so, we can incorporate different types of confidence intervals in the analysis of Monte Carlo simulation studies.
Compare with the default setting:

```{r}
s2 <- simsum(data = tt, estvarname = "diff", true = -1, se = "se", methodvar = "method", by = "dgm")
summary(s2, stats = "cover")
```

The `ci.limits` is also useful when using non-symmetrical confidence intervals, e.g. when using bootstrapped confidence intervals.

A pair of values can also be passed to `rsimsum` as the `ci.limits` argument:

```{r}
s3 <- simsum(data = tt, estvarname = "diff", true = -1, se = "se", ci.limits = c(-1.5, -0.5), methodvar = "method", by = "dgm")
summary(s3, stats = "cover")
```

If you have a better example of the utility of this method please [get in touch](https://twitter.com/ellessenne) - I'd love to hear from you!

Finally, we can pass a column of values for `true` as well:

```{r}
tt$true <- -1
s4 <- simsum(data = tt, estvarname = "diff", true = "true", se = "se", ci.limits = c("lower", "upper"), methodvar = "method", by = "dgm")
summary(s4, stats = "cover")
```

Compare with the default settings:

```{r}
summary(s2, stats = "cover")
```
